### 2022-01-25 src/text/0-authors.tex

\authorrunning{Deil, Donath, Terrier et al.}

%Opt in solution for the contributor list...

\author{
Christoph Deil\inst{\ref{inst:mpik}}
\and
Axel Donath\inst{\ref{inst:mpik}}
\and
Régis Terrier\inst{\ref{inst:apc}}
\and
Johannes King\inst{\ref{inst:mpik}}
\and
Léa Jouvin\inst{\ref{inst:apc}}
\and
Manuel Paz Arribas\inst{\ref{inst:humboldt}}
\and
Ellis Owen\inst{\ref{inst:ucl}}
\and
Julien Lefaucheur\inst{\ref{inst:obsparis}}
\and
Olga Vorokh\inst{\ref{inst:olga}}
\and
Jonathan Harris\inst{\ref{inst:jon}}
\and
Brigitta Sipocz\inst{\ref{inst:brigitta1}, \ref{inst:brigitta2}}
\and
Dirk Lennarz\inst{\ref{inst:atlanta}}
\and
Helen Poon\inst{\ref{inst:mpik}}
\and
Nachiketa Chakraborty\inst{\ref{inst:mpik}}
\and
Domenico Tiziani\inst{\ref{inst:erlangen}}
\and
Luigi Tibaldo\inst{\ref{inst:mpik}}
\and
Victor Zabalza\inst{\ref{inst:victor}}
\and
Rolf Bühler\inst{\ref{inst:desy}}
\and
Stefan Klepser\inst{\ref{inst:desy}}
\and
Ignasi Reichardt\inst{\ref{inst:padova}}
}

\institute{
Max-Planck-Institut f\"{u}r Kernphysik, Heidelberg, Germany
\label{inst:mpik}
\and
APC, University of Paris 7, France
\label{inst:apc}
\and
LUTH, Observatoire de Paris, Meudon, France
\label{inst:obsparis}
\and
DESY, Zeuthen, Germany
\label{inst:desy}
\and
Humboldt University, Berlin, Germany
\label{inst:humboldt}
\and
UCL-MSSL, Dorking, United Kingdom
\label{inst:ucl}
\and
INFN, Padova, Italy
\label{inst:padova}
\and
Belarusian State University, Belarus, Minsk
\label{inst:olga}
\and
FAU, Erlangen, Germany
\label{inst:erlangen}
\and
School of Physics and Center for Relativistic Astrophysics, Georgia
Institute of Technology, Atlanta, GA, USA
\label{inst:atlanta}
\and
Centre for Astrophysics Research, University of Hertfordshire, College Lane, Hatfield, AL10 9AB, UK
\label{inst:brigitta1}
\and
Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK
\label{inst:brigitta2}
\and
Department of Physics, Durham University, South Road, Durham, DH1 3LE, UK
\label{inst:jon}
\and
TODO: Victor address
\label{inst:victor}
}