### 2022-01-25 src/text/5-summary-outlook.tex

\section{Summary and Outlook}
\label{sec:so}

\todo{Axel and Regis write this...}

Summary what we have in v0.9 and presented in this paper.

Roadmap to v1.0, about half a page.

Short conclusion: Gammapy has potential to be the Python package for gamma-ray astronomy.

Prospects for HAWC / SWGO? Or speak in general about water Cherenkov observatories...