### 2022-01-25 src/text/6-acknowledgements.tex

\begin{acknowledgements}

Mention Christoph here?

We would like to thank the \texttt{Numpy}, \texttt{Scipy}, \texttt{IPython} and
\texttt{Matplotlib} communities for providing their packages which are invaluable
to the development of Astropy. We thank the GitHub team for providing us with
an excellent free development platform. We also are grateful to Read the Docs
(\ReadthedocsUrl), and Travis
(\TravisUrl) for providing free documentation
hosting and testing respectively. Finally, we would like to thank all the
Gammapy users that have provided feedback and submitted bug reports.

TODO: copy over stuff from \url{http://docs.gammapy.org/en/latest/about.html#thanks}.
TODO: add the ANR for Luca (and Atreyee in LUPM?)

\end{acknowledgements}