### 2022-01-27 src/text/0-authors.tex
Reformat tex files

\authorrunning{Deil, Donath, Terrier et al.}

%Opt in solution for the contributor list...

\author{
	Christoph Deil\inst{\ref{inst:mpik}} \and Axel Donath\inst{\ref{inst:mpik}}
	\and Régis Terrier\inst{\ref{inst:apc}} \and Johannes
	King\inst{\ref{inst:mpik}} \and Léa Jouvin\inst{\ref{inst:apc}} \and Manuel Paz
	Arribas\inst{\ref{inst:humboldt}} \and Ellis Owen\inst{\ref{inst:ucl}} \and
	Julien Lefaucheur\inst{\ref{inst:obsparis}} \and Olga
	Vorokh\inst{\ref{inst:olga}} \and Jonathan Harris\inst{\ref{inst:jon}} \and
	Brigitta Sipocz\inst{\ref{inst:brigitta1}, \ref{inst:brigitta2}} \and Dirk
	Lennarz\inst{\ref{inst:atlanta}} \and Helen Poon\inst{\ref{inst:mpik}} \and
	Nachiketa Chakraborty\inst{\ref{inst:mpik}} \and Domenico
	Tiziani\inst{\ref{inst:erlangen}} \and Luigi Tibaldo\inst{\ref{inst:mpik}} \and
	Victor Zabalza\inst{\ref{inst:victor}} \and Rolf Bühler\inst{\ref{inst:desy}}
	\and Stefan Klepser\inst{\ref{inst:desy}} \and Ignasi
	Reichardt\inst{\ref{inst:padova}} }

\institute{
Max-Planck-Institut f\"{u}r Kernphysik, Heidelberg, Germany \label{inst:mpik}
\and APC, University of Paris 7, France \label{inst:apc} \and LUTH,
Observatoire de Paris, Meudon, France \label{inst:obsparis} \and DESY, Zeuthen,
Germany \label{inst:desy} \and Humboldt University, Berlin, Germany
\label{inst:humboldt} \and UCL-MSSL, Dorking, United Kingdom \label{inst:ucl}
\and INFN, Padova, Italy \label{inst:padova} \and Belarusian State University,
Belarus, Minsk \label{inst:olga} \and FAU, Erlangen, Germany
\label{inst:erlangen} \and School of Physics and Center for Relativistic
Astrophysics, Georgia Institute of Technology, Atlanta, GA, USA
\label{inst:atlanta} \and Centre for Astrophysics Research, University of
Hertfordshire, College Lane, Hatfield, AL10 9AB, UK \label{inst:brigitta1} \and
Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3
0HA, UK \label{inst:brigitta2} \and Department of Physics, Durham University,
South Road, Durham, DH1 3LE, UK \label{inst:jon} \and TODO: Victor address
\label{inst:victor} }