### 2022-01-28 src/text/0-authors.tex
Adapt authors script

\authorrunning{Deil, Donath, Terrier et al.}

\author{
	Axel Donath \inst{\ref{inst:3}} \and
	Christoph Deil \inst{\ref{inst:8}} \and
	Régis Terrier \inst{\ref{inst:1}} \and
	Johannes King \inst{\ref{inst:15}} \and
	Jose Enrique Ruiz \inst{\ref{inst:11}} \and
	Quentin Remy \inst{\ref{inst:12}} \and
	Léa Jouvin \inst{\ref{inst:unknown}} \and
	Atreyee Sinha \inst{\ref{inst:7}} \and
	Matthew Wood \inst{\ref{inst:unknown}} \and
	Fabio Pintore \inst{\ref{inst:10}} \and
	Manuel Paz Arribas \inst{\ref{inst:unknown}} \and
	Laura Olivera \inst{\ref{inst:12}} \and
	Luca Giunti \inst{\ref{inst:0}} \and
	Bruno Khelifi \inst{\ref{inst:2}} \and
	Ellis Owen \inst{\ref{inst:unknown}} \and
	Brigitta Sipőcz \inst{\ref{inst:6}} \and
	Olga Vorokh \inst{\ref{inst:unknown}} \and
	Julien Lefaucheur \inst{\ref{inst:unknown}} \and
	Fabio Acero \inst{\ref{inst:4}} \and
	Thomas Robitaille \inst{\ref{inst:unknown}} \and
	David Fidalgo \inst{\ref{inst:unknown}} \and
	Jonathan D. Harris \inst{\ref{inst:unknown}} \and
	Lars Mohrmann \inst{\ref{inst:13}} \and
	Cosimo Nigro \inst{\ref{inst:9}} \and
	Dirk Lennarz \inst{\ref{inst:unknown}} \and
	Jalel Eddine Hajlaoui \inst{\ref{inst:unknown}} \and
	Alexis de Almeida Coutinho \inst{\ref{inst:unknown}} \and
	Matthias Wegenmat \inst{\ref{inst:unknown}} \and
	Dimitri Papadopoulos \inst{\ref{inst:unknown}} \and
	Maximilian Nöthe \inst{\ref{inst:unknown}} \and
	Nachiketa Chakraborty \inst{\ref{inst:unknown}} \and
	Michael Droettboom \inst{\ref{inst:14}} \and
	Jason Watson \inst{\ref{inst:5}} \and
	Helen Poon \inst{\ref{inst:unknown}} \and
	Arjun Voruganti \inst{\ref{inst:unknown}} \and
	Vikas Joshi \inst{\ref{inst:unknown}} \and
	Thomas Armstrong \inst{\ref{inst:unknown}} \and
	Erik Tollerud \inst{\ref{inst:16}} \and
	Erik M. Bray \inst{\ref{inst:unknown}} \and
	Domenico Tiziani \inst{\ref{inst:unknown}} \and
	Gabriel Emery \inst{\ref{inst:unknown}} \and
	Hubert Siejkowski \inst{\ref{inst:unknown}} \and
	Kai Brügge \inst{\ref{inst:unknown}} \and
	Luigi Tibaldo \inst{\ref{inst:unknown}} \and
	Arpit Gogia \inst{\ref{inst:unknown}} \and
	Ignacio Minaya \inst{\ref{inst:unknown}} \and
	Marion Spir-Jacob \inst{\ref{inst:unknown}} \and
	Yves Gallant \inst{\ref{inst:unknown}} \and
	Andrew W. Chen \inst{\ref{inst:unknown}} \and
	Roberta Zanin \inst{\ref{inst:unknown}} \and
	Jean-Philippe Lenain \inst{\ref{inst:unknown}} \and
	Larry Bradley \inst{\ref{inst:unknown}} \and
	Kaori Nakashima \inst{\ref{inst:unknown}} \and
	Anne Lemière \inst{\ref{inst:unknown}} \and
	Mathieu de Bony \inst{\ref{inst:unknown}} \and
	Matthew Craig \inst{\ref{inst:unknown}} \and
	Lab Saha \inst{\ref{inst:unknown}} \and
	Zé Vinicius \inst{\ref{inst:unknown}} \and
	Kyle Barbary \inst{\ref{inst:unknown}} \and
	Thomas Vuillaume \inst{\ref{inst:unknown}} \and
	Adam Ginsburg \inst{\ref{inst:unknown}} \and
	Daniel Morcuende \inst{\ref{inst:unknown}} \and
	José Luis Contreras \inst{\ref{inst:unknown}} \and
	Laura Vega Garcia \inst{\ref{inst:unknown}} \and
	Oscar Blanch Bigas \inst{\ref{inst:unknown}} \and
	Víctor Zabalza \inst{\ref{inst:unknown}} \and
	Wolfgang Kerzendorf \inst{\ref{inst:unknown}} \and
	Rolf Buehler \inst{\ref{inst:unknown}} \and
	Sebastian Panny \inst{\ref{inst:unknown}} \and
	Silvia Manconi \inst{\ref{inst:unknown}} \and
	Stefan Klepser \inst{\ref{inst:unknown}} \and
	Peter Deiml \inst{\ref{inst:unknown}} \and
	Johannes Buchner \inst{\ref{inst:unknown}} \and
	Hugo van Kemenade \inst{\ref{inst:unknown}} \and
	Eric O. Lebigot \inst{\ref{inst:unknown}} \and
	Benjamin Alan Weaver \inst{\ref{inst:unknown}} \and
	Debanjan Bose \inst{\ref{inst:unknown}} \and
	Rubén López-Coto \inst{\ref{inst:unknown}} \and
	Sam Carter \inst{\ref{inst:unknown}}
}

\institute{
	APC \label{inst:0} \and
	Astroparticle and Cosmology, CNRS, Université de Paris \label{inst:1} \and
	CNRS \label{inst:2} \and
	Center for Astrophysics | Harvard \& Smithonian \label{inst:3} \and
	DAp/AIM, CEA-Saclay/CNRS \label{inst:4} \and
	DESY \label{inst:5} \and
	DIRAC Institute, UW \label{inst:6} \and
	EMFTEL, UCM \label{inst:7} \and
	HeidelbergCement \label{inst:8} \and
	IFAE \label{inst:9} \and
	INAF/IASF Palermo \label{inst:10} \and
	Instituto de Astrofísica de Andalucía - CSIC \label{inst:11} \and
	MPIK \label{inst:12} \and
	MPIK, Heidelberg \label{inst:13} \and
	Mozilla \label{inst:14} \and
	Pamyra \label{inst:15} \and
	Space Telescope Science Institute \label{inst:16} \and
	unknown \label{inst:unknown}
}