### 2022-01-31 src/text/2-package-subsections/astro.tex
Introduce subfolder for package subsections

\subsection{gammapy.astro}
\label{ssec:gammapy-astro}
\todo{Axel Donath}
Dark matter models, source population modelling