### 2022-01-31 src/text/2-package-subsections/catalog.tex

\subsection{gammapy.catalog}
\label{ssec:gammapy-catalog}
\todo{Axel Donath}
Gamma-ray catalog access

\begin{figure}
	\import{code-examples/generated/}{gp_catalogs}
	\caption{Using gammapy.catalogs}
	\label{codeexample:data}
\end{figure}