### 2022-01-31 src/text/2-package-subsections/maps.tex

\subsection{gammapy.maps}
\label{ssec:gammapy-maps}
\todo{Laura Olivera-Nieto}
The \verb|gammapy.maps| sub-package provides classes for representing pixelized
data structures with at least two spatial dimensions representing a set of
coordinates or a region on a sphere. In addition it allows to handle and
arbitrary number of non-spatial data axes, such as time or energy.

It provides a uniform API for WCS, HEALPix and region based data structures.

\begin{figure}
	\import{code-examples/generated/}{gp_maps}
	\caption{Using gammapy.data to access DL3 level data with a DataStore}
	\label{codeexample:data}
\end{figure}