### 2022-01-31 src/text/2-package-subsections/modeling.tex

\subsection{gammapy.modeling}
\label{ssec:gammapy-modeling}
\todo{Quentin Remy}
Models and fitting

\begin{figure}
	\import{code-examples/generated/}{gp_models}
	\caption{Using gammapy.modeling.models}
	\label{fig*:minted:gp_models}
\end{figure}