### 2022-01-31 src/text/2-package.tex

\section{Analysis Workflow Overview}
\label{sec:analysis-workflow-overview}

\section{Gammapy package}
\label{sec:gammapy-package}

The \gammapy package is structured into multiple sub-packages which mostly
follow the stages in the data reduction workflow.

\subsection{Overview}
\label{ssec:overview}
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/data_flow.pdf}
	\caption{
		Gammapy sub-package structure and data analysis workflow. }
	\label{fig:data_flow}
\end{figure*}

Outline: * List typical analysis use cases * Can use from Python and Jupyter ->
show Figure with Jupyter notebook here. * Gammapy code structure * How Numpy
and Astropy is used

Figures: * Add a Figure showing dataflow in a typical application DL3 at the
top, spectrum, map, lightcurve, fit results at the bottom. Mention major
classes in between (DataStore, EventList, Map, MapMaker, MapFit, …) * Probably
not: Figure showing sub-packages and how they relate (gammapy.data and
gammapy.irf at the base, then gammapy.maps, etc. * The code example Figure how
to make a counts map, to explain how the package works.

\todo{How to sort the sub-packages? After data flow or alphabetically?}
\input{text/2-package-subsections/analysis}
\input{text/2-package-subsections/astro}
\input{text/2-package-subsections/catalog}
\input{text/2-package-subsections/data}
\input{text/2-package-subsections/datasets}
\input{text/2-package-subsections/estimators}
\input{text/2-package-subsections/irf}
\input{text/2-package-subsections/makers}
\input{text/2-package-subsections/modeling}
\input{text/2-package-subsections/stats}
\input{text/2-package-subsections/utils}
\input{text/2-package-subsections/visualisation}