### 2022-01-31 src/text/2-package-subsections/analysis.tex

\subsection{gammapy.analysis}
\label{ssec:gammapy-analysis}
\todo{Jose Enrique writes this...}
High level analysis API