### 2022-01-31 src/text/2-package-subsections/data.tex

\subsection{gammapy.data}
\label{ssec:gammapy-data}
\todo{Cosimo Nigro}
The \verb|gammapy.data| sub-package provides access to DL3 level data and
observation handling.

\begin{figure}

	\import{code-examples/generated/}{gp_data}

	\caption{Using gammapy.data to access DL3 level data with a DataStore}
	\label{fig*:minted:gp_data}
\end{figure}