### 2022-01-31 src/text/2-package-subsections/irf.tex

\subsection{gammapy.irf}
\label{ssec:gammapy-irf}
\todo{Fabio Pintore}
IRF classes