### 2022-01-31 src/text/2-package-subsections/visualisation.tex

\subsection{gammapy.visualisation}
\label{ssec:gammapy-visualisation}
\todo{Axel Donath}
Plotters etc.