### 2022-02-07 src/text/2-package-subsections/makers.tex
Add Makers text.

\subsection{gammapy.makers}
\label{ssec:gammapy-makers}
\todo{Regis Terrier}

The data reduction contains all tasks required to process and prepare data at the DL3
level for modeling and fitting. Events are binned and IRFs are interpolated and
projected onto the chosen analysis geometry. This produces counts, exposure, background, psf
and energy dispersion maps.
The {\tt gammapy.makers} sub-package contains the various classes and functions required.

Because the background models suffer from strong uncertainties it is required to correct
them from the data themselves. Several techniques are commonly used in gamma-ray astronomy
such as field-of-view background normalization or background measurement in reflected regions 
regions, see~\cite{test}. Specific {\tt Makers} such as the {\tt FoVBackgroundMaker} or the
{\tt ReflectedRegionsBackgroundMaker} are in charge of this step.

Finally, to limit other sources of systematic uncertainties, a data validity domain is
determined by the {\tt SafeMaskMaker}.

\begin{figure}
	\import{code-examples/generated/}{gp_makers}

	\caption{Using gammapy.data to access DL3 level data}
	\label{ig*:minted:gp_makers}
\end{figure}