### 2022-02-11 src/text/2-package-subsections/maps.tex
start drafting

\subsection{gammapy.maps}
\label{ssec:gammapy-maps}
\todo{Laura Olivera-Nieto}
The \verb|gammapy.maps| sub-package provides classes for representing data structures associated with a set of coordinates or a region on a sphere. In addition it allows to handle an arbitrary number of non-spatial data dimensions, such as time or energy. It is organized around three types of structures: geometries, sky-maps and map axes, which inherit from the base classes Geom, Map and MapAxis respectively.

The geometry defines the pixelization scheme and map boundaries. It also provides methods to transform between sky and pixel coordinates. Maps consist of a geometry instance together with a data array containing the corresponding map values. Map axes contain a sequence of ordered values which define bins on a given dimension, spatial or not. Map axes can have physical units attached to them, as well as non-linear step sizes.

The sub-package provides a uniform API for the FITS World Coordinate System (WCS)~\citep{Calabretta2002}, the HEALPix pixelization~\citep{Gorski2005} and region-based data structures:

% itemize because it helps me write, could also just be paragraphs...
\begin{itemize}
	\item \textbf{WCS: } The FITS WCS pixelization supports a different number of projections to represent celestial spherical coordinates in a regular rectangular grid.
	\item \textbf{HEALPix: } This pixelization provides a subdivision of a sphere in which each pixel covers the same surface area as every other pixel. As a consequence, however, pixel shapes are no longer rectangular, or regular. Gammapy provides limited support to HEALPix-based maps, relying in some cases on projections to a WCS grid.
	\item \textbf{Region geometries: } In this case, instead of a fine spatial grid on a given sky region, the spatial dimension is reduced to a single bin with an arbitrary shape, describing a region in the sky with that same shape. Typically they are is used together with a non-spatial dimension, for example an energy axis to represent how a quantity varies with energy on a given region of the sky.
\end{itemize}

\begin{figure}
	\import{code-examples/generated/}{gp_maps}
	\caption{Using gammapy.data to access DL3 level data with a DataStore}
	\label{codeexample:data}
\end{figure}