### 2022-02-18 src/text/3-applications.tex

\section{Applications}
\label{sec:applications}

Each application example is a notebook in the online material: We could have
one analysis as Python scripts instead of notebook in the online material. At
the start of this section, point to gammapy-paper repo on Github and say that
there’s a Binder where people can try the examples online.

TODO: mention other application examples (joint Crab paper, HESS validation
paper, HGPS, ...) here or in a subsection "other applications" at the end of
this section?

\subsection{Source detection}
\label{ssec:source-detection}

See Figure~\ref{fig:fermi_ts_map}.

Ref:~\citep{Stewart2009}

\subsection{Multi instrument analysis}
\label{ssec:multi-instrument-analysis}