### 2022-02-21 src/text/2-package-subsections/datasets.tex
write datasets api

\subsection{gammapy.datasets}
\label{ssec:gammapy-datasets}
\todo{Atreyee Sinha}

The end product of the data reduction process described in \ref{ssec:gammapy-makers} are a set of binned counts, background and IRF maps, at the DL4 level. The gammapy.datasets subpackage contains classes to bundle together binned data along with associated models and the likelihood, which provides an interface to the Fit class \ref{ssec:gammapy-modeling} for modeling and fitting purposes. Figure \ref{fig*:minted:gp_datasets} Depending upon the type of analysis (3D or only spectral), and the associated statistics (Cash, wstat or chi2), different types of Datasets are supported. The predicted counts are computed by convolution of the models with the associate IRFs. Fitting of precomputed flux points is enabled through FluxPointsDatasets. Multiple datasets of same or different types can be  bundled together in Datasets, where the likelihood from each constituent member is added, thus facilitating joint fitting across different observations, and even different instruments across different wavelengths. Datasets also provide functionalities for manipulating reduced data, eg: stacking, sub-grouping, plotting, etc. Users can also create their customized datasets for implementing modified likelihood methods.

\begin{figure}
	\import{code-examples/generated/}{gp_datasets}
	\caption{Using gammapy.datasets for modelling and fitting}
	\label{fig*:minted:gp_datasets}
\end{figure}