### 2022-03-01 src/text/2-package-subsections/estimators.tex
Run tex indentation

\subsection{gammapy.estimators}
\label{ssec:gammapy-estimators}
\todo{Axel Donath}
The \verb|gammapy.estimators| sub-module features methods to compute flux
points, light curves, flux maps, flux profiles from data.

The initial fine binning of \verb|MapDataset| is grouped into larger bins.

Internal representation with a reference spectral model and an array of
normalisation values given in energy, time and spatial bins.

No unfolding.

\begin{itemize}
	\item Regrouping of dataset bins in time, energy etc.
	\item Reference spectral model scaled in energy bin
	\item "Forward folding" / "Backward folding": but there is no
	      difference between the two, backwards folding is forward folding with one bin
	\item Likelihood profiles \item Uniform N-dimensional data structure \item
	      Uniform API for plotting etc. \item \verb|FluxMaps| and \verb|FluxPoints| \item
	      Serialisation to multiple formats, Astropy's \verb|Table| and
	      \verb|BinnedTimeSeries|
	\item Additional quantities for debugging, such as predicted counts, fit convergence,
	      sum of fit statistics \item \end{itemize}

\begin{figure}
	\import{code-examples/generated/}{gp_estimators}
	\caption{Using the TSMapEstimator from gammapy.estimators to compute a
		sqrt(TS) map.} \label{fig*:minted:gp_estimators} \end{figure}