### 2022-03-07 src/text/3-applications.tex
reorder applications sections

\section{Applications}
\label{sec:applications}

Show some typical and advanced science cases, and refer to published papers

\subsection{Computing Flux points and significance maps}
\label{ssec:1D-3D-analysis}
\input{3.1-std-analysis}

- standard 1D and 3D analysis

See Figure~\ref{fig:fermi_ts_map}.

Ref:~\citep{Stewart2009}

\subsection{Temporal analysis}
\label{ssec:temporal-analysis}
\input{3.2-temporal}
lightcurves and temporal models

\subsection{Multi instrument analysis}
\label{ssec:multi-instrument-analysis}
\todo{Cosimo Nigro}
\begin{figure}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/multi_instrument_analysis.pdf}
	\caption{A multi-instrument analysis of the Crab Nebula}
	\label{fig:multi_instrument_analysis} 
\end{figure*}

\subsection{Fitting using physical models}
\label{ssec:naima-models-analysis}
\todo{Cosimo}
\begin{figure*}[t]
(added section in the text requesting multi-insturment figure)

\end{figure}

\subsection{Population studies}
\label{ssec:population-studies}
\todo{Quentin Remy}

Talk about simulations here? Or open a new section