### 2022-03-07 src/text/3-applications.tex
reorder applications sections

\section{Applications}
\label{sec:applications}

Show some typical and advanced science cases, and refer to published papers

\subsection{Computing Flux points and significance maps}
\label{ssec:1D-3D-analysis}
\input{3.1-std-analysis}

- standard 1D and 3D analysis

See Figure~\ref{fig:fermi_ts_map}.

Ref:~\citep{Stewart2009}

\subsection{Temporal analysis}
\label{ssec:temporal-analysis}
\input{3.2-temporal}
lightcurves and temporal models

\subsection{Multi instrument analysis}
\label{ssec:multi-instrument-analysis}
\todo{Cosimo Nigro}

\subsection{Fitting using physical models}
\label{ssec:naima-models-analysis}
\todo{}

\subsection{Population studies}
\label{ssec:population-studies}
\todo{Quentin Remy}

Talk about simulations here? Or open a new section

\input{text/3.1-cta}

\input{text/3.3-fermi}