### 2022-03-09 src/text/3-applications-subsections/1D-analysis.tex
Adapt filenames and input commands

\subsection{Classical spectral analysis}
\label{ssec:1D-analysis}
\todo{Axel}
A classical spectral analysis using Reflected Regions. Showing flux points, reflected regions on the CTA galactic center data

\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/cta_galactic_center.pdf}
	\caption{CTA Galactic Center example}
	\label{fig:cta_galactic_center}
\end{figure*}