### 2022-03-09 src/text/3-applications-subsections/multi-instrument-analysis.tex

\subsection{Multi instrument analysis}
\label{ssec:multi-instrument-analysis}
\todo{Cosimo Nigro}

\begin{figure*}[t]
	\sidecaption
	\includegraphics[width=0.666\textwidth]{figures/multi_instrument_analysis.pdf}
	\caption{A multi-instrument analysis of the Crab Nebula}
	\label{fig:multi_instrument_analysis}
\end{figure*}