### 2022-03-09 src/text/3-applications-subsections/population-studies.tex

\subsection{Population studies}
\label{ssec:population-studies}
\todo{Quentin Remy}