### 2022-03-09 src/text/3-applications.tex

\section{Applications}
\label{sec:applications}

Gammapy can be used for a variety of science cases by different IACT experiments. (Refer to publications)
In this section, we show a non-exhaustive list of some typical analysis that can be performed.
In general from the \gammapy side there is not limitation on which kind of analysis can be
done with wich instrument, however in practice it is limited by the availability of public data.

\input{text/3-applications-subsections/1D-analysis}
\input{text/3-applications-subsections/3D-analysis}
\input{text/3-applications-subsections/temporal-analysis}
\input{text/3-applications-subsections/multi-instrument-analysis}
\input{text/3-applications-subsections/sed-modelling-analysis}
\input{text/3-applications-subsections/population-studies}