### 2022-03-09 src/text/2-package-subsections/catalog.tex
Fix pre commit hooks

\subsection{gammapy.catalog}
\label{ssec:gammapy-catalog}
\todo{Atreyee}
Gamma-ray catalog access

Comprehensive source catalogs are increasingly being provided by many high
energy astrophysics experiments. gammapy.catalog provides a convenient access
to some common gamma-ray catalogs. A global catalog table is provided, along
with source models flux points and light curves (if available) for individual
objects, which are internally created from the supplied FITS tables. This
module works independently from the rest of the package, and the required
catalogs are supplied in GAMMAPY\_DATA. Presently, catalogs from Fermi-LAT
(standard~\citep{3FGL, 4FGL} and high energy~\citep{2FHL, 3FHL}), H.E.S.S.
galactic plane survey~\citep{HGPS}, HAWC~\citep{2HWC, 3HWC} and gamma-cat
~\citep{gamma-cat}; an open source gamma ray catalog combining information from
multiple catalogs) are supported.

\begin{figure}
	\import{code-examples/generated/}{gp_catalogs}
	\caption{Using gammapy.catalogs: Accessing underlying model, flux points and
		lightcurve from the Fermi-LAT 4FGL catalog for the blazar PKS 2155-304}
	\label{codeexample:data} \end{figure}