### 2022-03-09 src/text/2-package-subsections/datasets.tex

\subsection{gammapy.datasets}
\label{ssec:gammapy-datasets}
\todo{Atreyee Sinha}

The end product of the data reduction process described in Section
\ref{ssec:gammapy-makers} are a set of binned counts, background and IRF maps,
at the DL4 level. The gammapy.datasets subpackage contains classes to bundle
together binned data along with associated models and the likelihood, which
provides an interface to the Fit class (Sec \ref{ssec:gammapy-modeling}) for
modeling and fitting purposes. Depending upon the type of analysis and the
associated statistics, different types of Datasets are supported. MapDataset is
used for 3D (spectral and morphological) fitting, and a 1D spectral fitting is
done using SpectrumDatastet. While the default statistics for both of these is
Cash, their corresponding OnOff versions are adapted for the case where the
background is measured from real off counts, and support wstat statistics. The
predicted counts are computed by convolution of the models with the associate
IRFs. Fitting of precomputed flux points is enabled through FluxPointsDatasets,
using chi2 statistics. Multiple datasets of same or different types can be
bundled together in Datasets (e.g., Figure \ref{fig*:minted:gp_datasets}),
where the likelihood from each constituent member is added, thus facilitating
joint fitting across different observations, and even different instruments
across different wavelengths. Datasets also provide functionalities for
manipulating reduced data, eg: stacking, sub-grouping, plotting, etc. Users can
also create their customized datasets for implementing modified likelihood
methods.

\begin{figure}
	\import{code-examples/generated/}{gp_datasets}
	\caption{A Datasets container with FluxPointsDataset and MapDataset}
	\label{fig*:minted:gp_datasets}
\end{figure}