### 2022-03-09 src/text/3-applications-subsections/1D-analysis.tex

\subsection{1D Analysis}
\label{ssec:1d-analysis}
\todo{Axel}
A classical spectral analysis using Reflected Regions. Showing flux points,
reflected regions on the CTA galactic center data

\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/cta_galactic_center.pdf}
	\caption{CTA Galactic Center example}
	\label{fig:cta_galactic_center}
\end{figure*}