### 2022-03-09 src/text/3-applications.tex

\section{Applications}
\label{sec:applications}

Gammapy can be used for a variety of science cases by different IACT
experiments. (Refer to publications) In this section, we show a non-exhaustive
list of some typical analysis that can be performed. In general from the
\gammapy side there is not limitation on which kind of analysis can be done
with which instrument, however in practice it is limited by the availability of
public data.

\input{text/3-applications-subsections/1D-analysis}
\input{text/3-applications-subsections/3D-analysis}
\input{text/3-applications-subsections/temporal-analysis}
\input{text/3-applications-subsections/multi-instrument-analysis}
\input{text/3-applications-subsections/sed-modelling-analysis}
\input{text/3-applications-subsections/population-studies}