### 2022-03-21 src/text/6-acknowledgements.tex

\begin{acknowledgements}

	Mention Christoph here?

	We would like to thank the \texttt{Numpy}, \texttt{Scipy}, \texttt{IPython} and
	\texttt{Matplotlib} communities for providing their packages which are
	invaluable to the development of Astropy. We thank the GitHub team for
	providing us with an excellent free development platform. We also are grateful
	to Read the Docs (\ReadthedocsUrl), and Travis (\TravisUrl) for providing free
	documentation hosting and testing respectively. Finally, we would like to thank
	all the Gammapy users that have provided feedback and submitted bug reports.

	TODO: copy over stuff from
	\url{http://docs.gammapy.org/en/latest/about.html#thanks}.

	TODO: add the ANR for Luca (and Atreyee in LUPM?)

	J.E. Ruiz acknowledges financial support from the State Agency for Research of the Spanish MCIU through the "Center of Excellence Severo Ochoa" award to the Instituto de Astrof\'isica de Andaluc\'ia (SEV-2017-0709)

\end{acknowledgements}