### 2022-03-23 src/text/2-package-subsections/stats.tex
[skip ci] Add counts statisitics

\subsection{gammapy.stats}
\label{ssec:gammapy-stats}
\todo{Regis Terrier}

The gammapy.stats subpackage contains the fit statistics and associated statistical estimators
that are commonly used in gamma-ray astronomy.
In general, gamma-ray observations count Poisson-distributed events at various sky positions,
and contain both signal and background events. Estimation of the number of signal events is done
through likelihood maximization. In Gammapy, the fit statistics are Poisson log-likelihood functions
normalized like chi-squares, i.e. they follow the expression $2 \times log L$, where $L$ is the
likelihood function used.
The statistic function used when the expected number of background events is known
is the \emph{Cash} statistic ~\citep{Cash}. It is used by datasets using background
templates such as the MapDataset. When the number of background events is unknown and
an off measurement where only background events are expected is used, the statistic function
is WStat. It is a profile log-likelihood statistic where background counts are
marginalized parameters. It is used by datasets containing off counts measurements
such as the SpectrumDatasetOnOff, used for classical spectral analysis.

To perform simple statistical estimations on counts measurements, CountsStatistic classes
encapsulate the aforementioned statistic functions to measure excess counts and estimate the
associated statistical significance, errors and upper limits. They perform maximum likelihood
ratio tests to estimate significance (the square root of the statistic difference) and compute
likelihood profiles to measure errors and upper limits. The following example allows to compute
the Li \& Ma significance~\citep{LiMa} of a set of measurements: