### 2022-03-28 src/text/2-package-subsections/analysis.tex
analysis submodule text

\subsection{gammapy.analysis}
\label{ssec:gammapy-analysis}
\todo{Quentin, Axel...}

The gammapy.analysis sub-module provides a  high-level interface for the most common use cases identified in gamma-ray analyses. The classes and methods included can be used in Python scripts, notebooks or as commands within IPython sessions. The high-level user interface can also be used to automatise workflows driven by parameters declared in a configuration file in YAML format. This way a full analysis can be executed via a single command line taking the configuration file as input.

The Analysis class has the responsability of orchestrating of the workflow defined in the configuration (AnalysisConfig) and triggering the execution of the AnalysisStep classes that defines the common use cases identified. These steps include the following: 
 observations selection from the DataStore,  data reduction, excess map computation, model fitting, flux-points estimation, and light-curves production. The structure of the Analysis class allows to define custom AnalysisStep,  so ones could design and configure arbitrarily complex wokflows.