### 2022-03-30 src/text/2-package-subsections/analysis.tex

\subsection{gammapy.analysis}
\label{ssec:gammapy-analysis}
\todo{Quentin, Axel...}
High level analysis API