### 2022-03-30 src/text/2-package.tex
Add IRF formulae

\section{Analysis Workflow Overview}
\label{sec:analysis-workflow-overview}

% This might not be the bast place compared to introduction
The data processing workflow in Gamma-ray astronomy is usually split in two sequences.
The first one deals with the data processing from camera measurement, calibration, event
reconstruction and selection to yield a list of reconstructed gamma-ray event candidates.
This sequence, sometimes referred to as low-level analysis, is usually very specific to
a given observation technique if not a given instrument.

The other sequence, referred to as high-level analysis, deals with the extraction of physical
quantities related to gamma-ray sources and the production of high-level products such as spectra,
lightcurves and catalogs. The methods and tools applied here are more generic and are broadly
shared across the field. They also frequently imply joint analysis of multi-instrument data.
To extract physically relevant information, the measured data are usually compared to a
model of the expected gamma-ray emitters in the instrument field-of-view using statistical
techniques such as maximum likelihood.

We can write the expected number of detected events at measured position $p$ and energy $E$:
\begin{align}
   N(p, E) {\rm d}p {\rm d}E = &t_{\rm obs} \int_{E_{\rm true}} \int_{p_{\rm true}}  R(p, E|p_{\rm true}, E_{\rm true})\\
   &\times \Phi(p_{\rm true}, E_{\rm true}) {\rm d}E_{\rm true} {\rm d}p_{\rm true}
\end{align}
where
\begin{itemize}
\item $R(p, E| p_{\rm true}, E_{\rm true})$ is the instrument response
\item $\Phi(p_{\rm true}, E_{\rm true})$ is the sky flux model
\item $t_{\rm obs}$ is the observation time
\end{itemize}

A common assumption is that the instrument response can be simplified as the product
of three independent functions:

\begin{align}
   R(p, E|p_{\rm true}, E_{\rm true}) = &A_{\rm eff}(p_{\rm true}, E_{\rm true}) \times\\
    &PSF(p|p_{\rm true}, E_{\rm true}) \times\\
    &E_{\rm disp}(E|p_{\rm true}, E_{\rm true})
\end{align}
where:
\begin{itemize}
\item $A_{\rm eff}(p_{\rm true}, E_{\rm true})$ is the effective collection area of the detector. It is the product
  of the detector collection area times its detection efficiency at true energy $E_{\rm true}$ and position $p_{\rm true}$.
\item $PSF(p|p_{\rm true}, E_{\rm true})$ is the point spread function. It gives the probability of
  measuring a direction $p$ when the true direction is $p_{\rm true}$ and the true energy is $E_{\rm true}$.
  Gamma-ray instruments consider the probability density of the angular separation between true and reconstructed directions
  $\delta p = p_{\rm true} - p$, i.e. $PSF(\delta p|p_{\rm true}, E_{\rm true})$.
\item $E_{\rm disp}(E|p_{\rm true}, E_{\rm true})$ is the energy dispersion. It gives the probability to
  reconstruct the photon at energy $E$ when the true energy is $E_{\rm true}$ and the true position :$p_{\rm true}$.
  Gamma-ray instruments consider the probability density of the migration $\mu=\frac{E}{E_{\rm true}}$,
  i.e. $E_{\rm disp}(\mu|p_{\rm true}, E_{\rm true})$.
\end{itemize}

Gamma-ray data at the Data Level 3 therefore consists in lists of gamma-like events and their
corresponding instrument response functions (IRFs). The latter include the aforementioned
effective area, point spread function (PSF), energy dispersion and residual hadronic background.
The handling of DL3 data is performed by classes and methods
in the gammapy.data  (see \ref{ssec:gammapy-data}) and the gammapy.irf
(see \ref{ssec:gammapy-irf}) subpackages.

The first step in the analysis is the selection and extraction of observations
based of their meta data including information such as pointing direction, observation
time and observation conditions.

The next step of the analysis is the data reduction on a user-defined geometry.
A typical geometry consists in a spectral representation with a measured
energy axis, and in a spatial representation, either a coordinates system  with a projection
(for 3-dimensional or cube analysis) or a region on the sky (for regular spectral analysis).
The gammapy.maps subpackage provides general multidimensional geometry objects
(Geom) and the associated data structures (Maps), see \ref{ssec:gammapy-maps}.

All observation events and instrument responses are projected onto the
target geometry. Because residual hadronic background models can be subject
to significant uncertainties background correction must be applied,
such as the ring or the fied-of-view background techniques or
background measurements must be performed within, e.g. reflected regions~\citep{Berge07}.
Parts of the data with high associated IRF systematics must also be excluded by defining
a "safe" data range. These data reduction steps are performed by classes and functions
implemented in the gammapy.makers subpackage (see \ref{ssec:gammapy-makers}).

The counts data and the reduced IRFs in the form of maps are bundled into dataset objects
that represent the data level 4 (DL4). They can be written to
disk, in a format specific to Gammapy to allow users to read them back at any time later
for modeling and fitting.

This latter step
 datasets classes bundle reduced data in form of maps, reduced IRFs, models and
fit statistics. Different sub-classes support different analysis methods
and fit statistics (e.g. Poisson statistics with known background or
with OFF background measurements). The datasets are used to perform joint-likelihood
fitting allowing to combine different measurements, e.g. from different observations
but also from different instruments or event classes. They can also be used for binned
simulation as well as event sampling to simulate DL3 events data.

The next step is then typically to model and fit the datasets, either
individually, or in a joint likelihood analysis. For this purpose Gammapy
provides a uniform interface to multiple fitting backends. It also provides
a variety of :ref:`built in models <model-gallery>`. This includes spectral,
spatial and temporal model classes to describe the gamma-ray emission in the sky.
Independently or subsequently to the global modelling, the data can be
re-grouped to compute flux points, light curves and flux as well as significance
maps in energy bands.

\section{Gammapy package}
\label{sec:gammapy-package}

The \gammapy package is structured into multiple sub-packages which mostly
follow the stages in the data reduction workflow.

\subsection{Overview}
\label{ssec:overview}
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/data_flow.pdf}
	\caption{
		Gammapy sub-package structure and data analysis workflow. }
	\label{fig:data_flow} \end{figure*}

Outline: * List typical analysis use cases * Can use from Python and Jupyter ->
show Figure with Jupyter notebook here. * Gammapy code structure * How Numpy
and Astropy is used

Figures: * Add a Figure showing dataflow in a typical application DL3 at the
top, spectrum, map, lightcurve, fit results at the bottom. Mention major
classes in between (DataStore, EventList, Map, MapMaker, MapFit, …) * Probably
not: Figure showing sub-packages and how they relate (gammapy.data and
gammapy.irf at the base, then gammapy.maps, etc. * The code example Figure how
to make a counts map, to explain how the package works.

\todo{How to sort the sub-packages? After data flow or alphabetically? What
	about maps?}
\input{text/2-package-subsections/data}
\input{text/2-package-subsections/irf}
\input{text/2-package-subsections/maps}
\input{text/2-package-subsections/makers}
\input{text/2-package-subsections/datasets}
\input{text/2-package-subsections/modeling}
\input{text/2-package-subsections/stats}
\input{text/2-package-subsections/estimators}
\input{text/2-package-subsections/analysis}
\input{text/2-package-subsections/visualisation}
\input{text/2-package-subsections/astro}
\input{text/2-package-subsections/catalog}
\input{text/2-package-subsections/utils}