### 2022-04-04 src/text/3-applications-subsections/temporal-analysis.tex
write temporal section

\subsection{Temporal analysis}
\label{ssec:temporal-analysis}
\todo{Atreyee}

A common use case in most astrophysical scenarios is to study the temporal variability of the source. The most basic way to do this is to construct a lightcurve, ie, the flux of the source in each given time bin. In gammapy, this is done by using the `LightCurveEstimator` which fits the normalisation of the source in each energy band per observation. For custom time binning, an observation needs to be split into finer ones using the `Observation.select_time` method. Figure \ref{fig:hess_lightcurve_pks} shows the lightcurve of the blazar PKS 2155-304 in different energy bands as observed by the H.E.S.S. telescope during an exceptional flare on the night of July 29 - 30, 2006 \cite{2009A&A...502..749A}. The data is available publicly as a part of the HESS-DL3-DR1 \cite{HESS-DL3-DR1}, and shipped with GAMMAPY_DATA. Each observation is first split into 10 min smaller observations, and spectra extracted for each of these within a 0.11 deg radius around the source. A PowerLawSpectralModel with an index of 3.4 is assumed for all the datasets, and the `LightCurveEstimator` run for two energy bands, 0.5 - 1.5 TeV, and 1.5 - 20 TeV, resp.

Lightcurve in two energy bands (600 GeV - 1.5 TeV, and 1.5 TeV to 20 TeV) for
PKS2155-304 flare seen by H.E.S.S. as in our notebooks.

\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/hess_lightcurve_pks.pdf}
	\caption{10 min binned lightcurve for PKS~2155-304 in two energy bands, (500 GeV - 1.5 TeV, and 1.5 TeV to 20 TeV) as observed by the H.E.S.S. telescopes in 2006.}
	\label{fig:hess_lightcurve_pks} 
\end{figure*}

Instead of extracting a lightcurve, it is also possible to directly fit temporal models to the reduced datasets. By associating an appropriate SkyModel, consisting of both temporal and spectral components, to each dataset, a joint fitting across the datasets will directly return the best fit temporal and spectral parameters.