### 2022-04-08 src/text/4-project.tex

\section{Gammapy project}
\label{sec:gammapy-project}

In this section, we provide an overview of the organization of the Gammapy project. We briefly describe the main roles and responsibilities within the team, as well as the technical infrastructure designed to facilitate the development and maintenance of Gammapy as a high quality software. We use common tools and services for software development of Python open-source projects, code review, testing, package distribution and user support, with a customized solution for a versioned thoroughly tested documentation in the form of user-friendly playable tutorials. This section concludes with an outlook on the roadmap for future directions.

\subsection{Organizational structure}
\label{ssec:organizational-structure}

Gammapy is a well-organised international open-source project with a broad developer base and the presence of commitments from strong groups and institutes in the very high-energies astrophysics domain\footnote{\url{https://gammapy.org/team.html}}. The main development roadmaps are discussed and validated by a Coordination Committee, composed of representatives of the main contributing laboratories. This committee is chaired by a Project Manager and his deputy while two Lead Developers manage the development strategy and organise technical activities. Because of this institutionally driven characteristic, the permanent staff and commitment of supporting institutes ensure the continuity of the executive teams. A core team of developers from the contributing laboratories is in charge of the regular development, which benefits from punctual contributions of the community at large.

\subsection{Technical infrastructure}
\label{ssec:technical-infrastructure}

Gammapy follows an open-source and open-contribution development model based on the cloud repository service GitHub. A GitHub organization \textit{gammapy}\footnote{\url{https://github.com/gammapy}} hosts different repositories related with the project. The software codebase may be found in the \textit{gammapy} repository (see Table~\ref{table:codestats:data} and Figure~\ref{fig:codestats:lang} for code lines statistics). We make extensive use of the pull request system to discuss and review code contributions.

\begin{table}
	\import{tables/generated/}{codestats}
	\caption{Coding languages statistics in Gammapy project}
	\label{table:codestats:data}
\end{table}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.5\textwidth]{figures/codestats.pdf}
	\caption{
		Percentage of lines of code in Gammapy project} \label{fig:codestats:lang}
\end{figure}

Several automatized tasks are set as GitHub actions\footnote{\url{https://github.com/features/actions}}, blocking the processes and alerting developers when fails occur. This is the case of the content integration workflow, which monitors the execution of the test coverage suite\footnote{\url{https://pytest.org}} using datasets from the \textit{gammapy-data} repository. Tests scan not only the codebase but also the code snippets present in docstrings of the scripts and in the RST documentation files, as well as in the tutorials provided in the form of Jupyter notebooks.

Other automatized tasks, executing in the \textit{gammapy-benchmarks} repository, are responsible for numerical validation tests and benchmarks monitoring. Also tasks related with the release process are partially automatized, and every contribution to the codebase repository triggers the documentation building and publishing workflow within the \textit{gammapy-docs} repository (see Sec. ~\ref{ssec:software-distribution} and Sec. ~\ref{ssec:documentation}).

This small ecosystem of interconnected up-to-date repositories, automatized tasks and alerts, is just a part of a bigger set of GitHub repositories, where most of them are related with the project but not necessary for the development of the software (i.e. project webpage, complementary high-energy astrophysics object catalogs, coding sprints and weekly developer calls minutes, contributions to conferences, other digital assets, etc.) Finally, third-party services for code quality metrics are also set and may be found as status shields in the codebase repository.

\subsection{Software distribution}
\label{ssec:software-distribution}

Gammapy is distributed for Linux, Windows and Mac environments, and installed in the usual way for Python packages. Each stable release is uploaded to the Python package index\footnote{\url{https://pypi.org}} and as a binary package to the \textit{conda-forge} and \textit{astropy} Anaconda repository\footnote{\url{https://anaconda.org/anaconda/repo}} channels. At this time, Gammapy is also available as a Debian Linux package\footnote{\url{https://packages.debian.org/sid/python3-gammapy}}. We recommend installing the software using the \textit{conda} installation process with an environment definition file that we provide, so to work within a virtual isolated environment with additional useful packages and ensure reproducibility.

Gammapy is indexed in Astronomy Source Code Library\footnote{\url{https://ascl.net/1711.014}} and Zenodo\footnote{\url{https://zenodo.org/record/5721467o}} digital libraries for software. The Zenodo record is synchronised with the codebase GitHub repository so that every release triggers the update of the versioned record. In addition, the next release of Gammapy will be added to the Open-source scientific Software and Service Repository\footnote{\url{https://projectescape.eu/ossr}} and indexed in the European Open Science Cloud catalog\footnote{\url{https://eosc-portal.eu}}.

\subsection{Documentation and user-support}
\label{ssec:documentation-and-user-support}
Gammapy provides its user community with a tested and versioned up-to-date on-line documentation\footnote{\url{https://docs.gammapy.org}}~\citep{2019ASPC..523..357B} built with Sphinx\footnote{\url{https://www.sphinx-doc.org}} scanning the codebase Python scripts, as well as a set of RST files and Jupyter notebooks. The documentation includes a handwritten user guide, a set of executable tutorials, and a reference to the API automatically extracted from the code and docstrings. The Gammapy code snippets present in the documentation are tested in different environments using our continuous integration (CI) workflow based on GitHub actions.

The Jupyter notebooks tutorials are stored stripped of their output cells, which greatly helps in identifying differences in the contribution review. During the CI these are executed, the output cells parsed and, in case an error is found in one of the output cells then validation fails. Code formatting of the input cells is also done, parsing their content and using a particular syntax to provide links to the API documentation pages. The resulting web published tutorials also provide links to playground spaces in myBinder~\citep{project_jupyter-proc-scipy-2018}, where they may be executed on-line in versioned virtual environments hosted in the myBinder infrastructure. Users may also play with the tutorials locally in their laptops. They can download a specific version of the tutorials together with the associated datasets needed and the specific conda computing environment, using the \textit{gammapy download} command.

We have also set up a solution for users to share recipes as Jupyter notebooks that do not fit in the Gammapy core documentation but which may be relevant as specific use cases. Contributions happen via pull requests to the \textit{gammapy-recipes} GitHub repository and merged after a short review. All notebooks in the repository are tested and published in the Gammapy recipes webpage\footnote{\url{https://gammapy.github.io/gammapy-recipes}} automatically using GitHub actions.

A growing community of users is gathering around the Slack messaging\footnote{\url{https://gammapy.slack.com}} and GitHub discussions\footnote{\url{https://github.com/gammapy/gammapy/discussions}} support forums, providing valuable feedback on the Gammapy functionalities, interface and documentation. Other communication channels have been set like mailing lists, a Twitter account\footnote{\url{https://twitter.com/gammapyST}}, regular public coding sprint meetings, hands-on session within collaborations, weekly development meetings, etc.

\subsection{Roadmap}
\todo{Axel, Regis}

\begin{itemize}
	\item pigs
	\item roadmap
\end{itemize}