### 2022-05-02 src/text/3-applications-subsections/multi-instrument-analysis.tex
Reformat text files

\subsection{Multi instrument analysis}
\label{ssec:multi-instrument-analysis}

\begin{figure*}[t]
	\sidecaption
	\includegraphics[width=0.666\textwidth]{figures/multi_instrument_analysis.pdf}
	\caption{A multi-instrument analysis of the Crab Nebula}
	\label{fig:multi_instrument_analysis}
\end{figure*}

%% should we mention the joint-crab here or in the intro, where we talk about
%% DL3, multi-instrument analysis and so forth...
%% cite Laura's DL3 HAWC paper with the updated version of the joint-crab spectrum?

In this multi-instrument analysis example we showcase the capabilities of
\gammapy to perform a simultaneous likelihood fit incorporating data from
different instruments and at different levels of reduction. We estimate the
spectrum of the Crab Nebula combining data from \fermi, \magic and \hawc. Maps
of \fermi data are prepared selecting a region of $X^{\circ}$ around the
position of the Crab Nebula applying the same selection criteria of the 3FHL
catalog (7 years of data with energy from $10\,{\rm GeV}$ to $2\,{\rm TeV}$,
\citealt{3FHL}).  The \magic data are two%% shall we explain exactly which type of Map is created?
observations of $20\,{\rm min}$ each, chosen from the dataset used to estimate
the performance of the upgraded stereo system \citep{magic_performance} and
already included in \cite{joint_crab}. The observations were taken at small
zenith angles ($<30^{\circ}$) in wobble mode \citep{fomin_1994}, with the
source sitting at an offset of $0.4^{\circ}$ from the FoV center. Their energy
range spans $80\,{\rm GeV} -- 20\,{\rm TeV}$. They are reduced to ON/OFF
dataset before being fitted. \hawc flux points data are estimated in
\cite{hawc_crab_2019} with $2.5\,{\rm years}$ of data and span an energy range
$300\,{\rm GeV} -- 300\,{\rm TeV}$, and directly read with \gammapy.

\gammapy automatically generates a likelihood including three different types
of terms. Two poissonian likelihoods: one for the \fermi map and one for the
ON/OFF counts, and a $\chi^2$ accounting for the flux points. For \fermi, a a
three-dimensional forward folding of the sky model with the IRF is performed,
in order to compute the predicted counts in each sky-coordinate and energy bin.
For \magic, a one-dimensional forward folding of the spectral model with the
IRF is performed to predict the counts in each estimated energy bin. A log
parabola is fitted to the almost fived decades in energy $10\,{\rm GeV} --
	300\,{\rm TeV}$. %% is the LP formula specified elsewhere?

The result of the joint fit is displayed in
Fig.~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative, we display the flexibility of \gammapy in
simultaneously fitting multi-isntrument data even at different levels of
reduction, we do not aim to provide a new measurement of the Crab Nebula
spectrum.

%% eventual part for physical modelling with naima