### 2022-05-02 src/text/3-applications-subsections/population-studies.tex

\subsection{Surveys, catalogs, and population studies}
\label{ssec:surveys-catalogs-and-population-studies}

Sky surveys have a large potential for new source detections, and new phenomena
discovery. They also offer less selection bias to perform source population
studies over a large set of coherently detected and modelled objects. Early
versions of \textit{gammapy} were developed in parallel of the preparation of
the HESS Galactic plane survey catalog~\citep[HGPS][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies~\citep{2018A&A...612A...2H,
	2018A&A...612A...3H}.

The increase in sensitivity  and resolution provided by new generation of
instruments scale up the number of sources detectable and the complexity of the
models needed to represent them accurately. As an example If we compare the
results of the HGPS to the expectations from the CTA Galactic Plane survey
simulations we jump from 78 sources detected by HESS to about 500 detectable by
CTA~\citep{2021arXiv210903729R}.

Studies performed on simulations not only offer a first glimpse on what could
be the sky seen by CTA (according to our current knowledge on source
populations), but also give us the opportunity to test the software on complex
use cases\footnote{Note that the CTA-GPS simulations were performed with the
	\textit{ctools} package~\citep{2016A&A...593A...1K} and analysed with both
	\textit{ctools} and \textit{gammapy} packages in order to cross-validate
	them.}. So we can  improve performances, optimize our analyses strategies, and
identify the needs in term of parallelisation to process the large datasets
provided by the surveys.

In short the production of catalogs from gamma-ray surveys can be divided in
four main steps : data-reduction; object detection; model fitting and model
selection; associations and classification. The IACTs data-reduction step is
done in the same way than described in the previous sections but scale-up to
few thousand of observations. The object detection step consists a minima in
finding local maxima in the significance, or TS maps,  given by the
ExcessMapEstimator, or TSMapEstimator, respectively.  Further refinements can
include for example  filtering and detection on these maps with techniques from
scikit-image package~\citep{scikit-image}, and outlier detection from
scikit-learn package~\citep{scikit-learn}. Tests on simulations shown that it
reduces the spurious detections at this stage and then speed up the next step
as less objects will have to be fitted simultaneously. During the modelling
step each object is alternatively fitted with different models in order to
determine their optimal parameters, and the best-candidate model. The
subpackage gammapy.modeling.models offers a large variety of choice, and the
possibility to add custom models.  Several spatial models (point-source, disk,
gaussian...), and spectral models (power-law, log-parabola...) may be tested
for each object, so the complexity of the problem increases rapidly in regions
crowded with multiple extended sources. Finally an object is discarded if its
best-fit model is not significantly preferred over the null hypothesis (no
source) comparing the difference in log-Likelihood between these two
hypotheses. For the association and classification step, that is tightly
connected to the population studies, we can a minima compare the fitted models
to the set of gammapy-ray catalogs available in gammapy.catalog. However
further multi-wavelength cross-matches are usually required to characterize the
sources.