### 2022-06-07 src/text/2-package-subsections/maps.tex
minor changes to maps

\subsection{gammapy.maps}
\label{ssec:gammapy-maps}
\todo{Laura Olivera-Nieto}
The \verb|gammapy.maps| sub-package provides classes that represent data
structures associated with a set of coordinates or a region on a sphere. In
addition it allows to handle an arbitrary number of non-spatial data
dimensions, such as time or energy. It is organized around three types of
structures: geometries, sky-maps and map axes, which inherit from the base
classes Geom, Map and MapAxis respectively.

The geometry object defines the pixelization scheme and map boundaries. It also
provides methods to transform between sky and pixel coordinates. Maps consist
of a geometry instance together with a data array containing the corresponding
map values. Map axes contain a sequence of ordered values which define bins on
a given dimension, spatial or not. Map axes can have physical units attached to
them, as well as non-linear step sizes.

The sub-package provides a uniform API for the FITS World Coordinate System
(WCS)~\citep{Calabretta2002}, the HEALPix pixelization~\citep{Gorski2005} and
region-based data structure (see Figure~\cite{ig*:minted:gp_maps}).

% itemize because it helps me write, could also just be paragraphs...
\begin{itemize}
	\item \textbf{WCS: } The FITS WCS pixelization supports a different
	      number of projections to represent celestial spherical coordinates in a regular
	      rectangular grid. Gammapy provides full support to data structures using this
	      pixelization scheme.
	 \item \textbf{HEALPix: } This pixelization provides a
	      subdivision of a sphere in which each pixel covers the same surface area as
	      every other pixel. As a consequence, however, pixel shapes are no longer
	      rectangular, or regular. \gammapy provides limited support to HEALPix-based
	      maps, relying in some cases on projections to a local WCS grid. 
   \item \textbf{Region geometries: } In this case, instead of a fine spatial grid
	      dividing a rectangular sky region, the spatial dimension is reduced to a single
	      bin with an arbitrary shape, describing a region in the sky with that same
	      shape. Typically they are is used together with a non-spatial dimension, for
	      example an energy axis, to represent how a quantity varies in that dimension
	      inside the corresponding region. \end{itemize}

Additionally, the MapAxis class provides a uniform API for axes representing
bins on any physical quantity, such as energy or angular offset. The special
case of time is covered by the dedicated TimeMapAxis, which allows time bins to
be non-contiguous, as it is often the case with observation time-stamps. The
generic class LabelMapAxis allows the creation of axes for non-numeric entries.

\begin{figure}
	\import{code-examples/generated/}{gp_maps}

	\caption{Using gammapy.maps to create a WCS, a HEALPix and a region
		map. Note the uniform API for map creation.} \label{ig*:minted:gp_maps}
\end{figure}