### 2022-07-08 src/text/2-package-subsections/irf.tex

\subsection{gammapy.irf}
\label{ssec:gammapy-irf}
\todo{Fabio Pintore}

\begin{figure}
	%	\import{code-examples/generated/}{gp_makers}

	\caption{In Fig.~\ref{ig*:minted:irf_examples}, we show some example of Aeff,
		PSF, Edisp and BKG read and plotted from a typical \irf.}
	\label{ig*:minted:irf_examples}
\end{figure}

The \code{gammapy.irf} sub-package contains all classes and functionality
to handle IRFs in a variety of formats.
Usually, \irfs store instrument properties in the form of multi-dimension
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles, detector coordinates and so on. The main information stored in
the common \gammaray \irfs are the effective area (Aeff), energy dispersion
(Edisp), point spread function (PSF) and background (BKG). The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units in different kinds, plot or write them into
output files. In the following, we list the main sub-classes of the
sub-package:

\paragraph{Effective area:}
\gammapy provides the class \code{EffectiveAreaTable2D} to
manage Aeff, which is usually defined in terms of true energy and offset angle.
The class functionalities offer the possibility to read from files or to create
it from scratch. {\code{EffectiveAreaTable2D} can also convert, interpolate,
write, and evaluate the effective area for a given energy and offset angles, or
even plot the multi-dimensional Aeff table.

\paragraph{Point spread function:}
\gammapy allows user to treat different kinds of PSFs,
in particular, parametric multi-dimensional gaussians (\code{EnergyDependentMultiGaussPSF})
or King profile functions (\code{PSFKing})
one. The \code{EnergyDependentMultiGaussPSF} class is able to handle up to three
gaussians, defined in terms of amplitudes and sigma given for each true energy
and offset angle bin. Similarly, \code{PSFKing takes} into account the gamma and
sigma parameters as defined here. The \code{ParametricPSF} allows to create a
PSF with a representation different from gaussian(s) or King profile(s).
Finally, the user can take advantage from the \code{PSFMap} class, which creates
a multi-dimensional map of the PSF in WCS coordinates. At each position, a PSF
kernel map (\code{PSFKernel}) provides a PSF as a function of the true energy.
The creation of PSF kernel maps, where the PSF is defined for each
sky-position, is also given to the user. The latter two can speed up analyses.

\paragraph{Energy dispersion:}
Edisp, in \iact, is generally given in terms of the so
called migration parameter ($\mu$), which is defined as the ratio between the
reconstructed energy and the true energy. This ratio should be as close as one
and its dispersion can assume the shape of a gaussian (or even more complex
distributions). Migration parameter is given at each offset angle and
reconstructed energy. The main sub-classes are \code{EnergyDispersion2D},
designed to interpret Edisp, \code{EDispKernelMap}, which builds an Edisp kernel
map, i.e., a 4-dimensional WCS map where at each sky-position is associated an
Edisp kernel. The latter is a representation of the Edisp as a function of the
true energy only thanks to the sub-class \code{EDispKernel}.

\paragraph{Background model:}
The BKG can be represented in \gammapy as either 1) a 2D map (\code{Background2D})
of count rate normalised per steradians and energy at different
energies and offset-angles or 2) as rate per steradians and energy, as a
function of reconstructed energy and detector coordinates (\code{Background3D}).
In the former, the background is expected to follow a radially symmetric shape,
while in the latter, it can be more complex.