### 2022-07-08 src/text/2-package-subsections/catalog.tex

\subsection{gammapy.catalog}
\label{ssec:gammapy-catalog}
\todo{Atreyee}
Gamma-ray catalog access

Comprehensive source catalogs are increasingly being provided by many high
energy astrophysics experiments. gammapy.catalog provides a convenient access
to some common gamma-ray catalogs. A global catalog table is provided, along
with source models flux points and light curves (if available) for individual
objects, which are internally created from the supplied FITS tables. This
module works independently from the rest of the package, and the required
catalogs are supplied in GAMMAPY\_DATA. The overview of currenly supported
catalogs and the correponding \gammapy classes is shown in Table~\ref{tab:catalogs}.
Newly released relevant catalogs will be added in future.

\begin{table*}
    \begin{center}
        \begin{tabular}{llll}
         \hline
         Class & Shortcut & Description & Reference\\
         \hline
         \code{SourceCatalog3FGL} & \code{"3fgl"} & Third catalog of \fermi sources & \cite{3FGL} \\
         \code{SourceCatalog4FGL} & \code{"4fgl"} & Fourth catalog of \fermi  sources & \cite{4FGL} \\
         \code{SourceCatalog2FHL} & \code{"2fhl"} & Second catalog high energy \fermi  sources & \cite{2FHL} \\
         \code{SourceCatalog3FHL} & \code{"3fhl"} & Third catalog high energy \fermi  sources & \cite{3FHL} \\
         \code{SourceCatalog2HWC} & \code{"2hwc"} & Second catalog of \hawc sources & \cite{2HWC} \\
         \code{SourceCatalog3HWC} & \code{"3hwc"} & Third catalog of \hawc sources & \cite{3HWC} \\
         \code{SourceCatalogHGPS} & \code{"hgps"} & \hess Galactic Plane Survey catalog & \cite{HGPS} \\
         \code{SourceCatalogGammaCat} & \code{"gammacat"} & Open source data collection & \citep{gamma-cat} \\
         \hline
         \end{tabular}
    \end{center}
    \label{tab:catalogs}
    \caption{Built-in catalogs.}
\end{table*}

\begin{figure}
	\import{code-examples/generated/}{gp_catalogs}
	\caption{Using gammapy.catalogs: Accessing underlying model, flux points and
		lightcurve from the Fermi-LAT 4FGL catalog for the blazar PKS 2155-304}
	\label{codeexample:data}
\end{figure}