### 2022-07-08 src/text/2-package-subsections/data.tex

\subsection{gammapy.data}
\label{ssec:gammapy-data}
\todo{Cosimo Nigro}

\begin{figure}

	\import{code-examples/generated/}{gp_data}

	\caption{Using \code{gammapy.data} to access DL3 level data with a \code{DataStore} object.
    Individual observations can be accessed by their unique integer observation id number.
    The actual events and instrument response functions can be accessed
    as attributes on the \code{Observation} object, such as \code{.events}
    or \code{.aeff} for the effective area information.
    }
	\label{fig*:minted:gp_data}
\end{figure}

As illustrated in Fig.~\ref{fig:data_flow}, the \code{gammapy.data} sub-package
implements the interface to the lowest data level, providing the input interface to gamma-ray
data. The \gammapy data model follows the \gadf specifications
(Sec.~\ref{sec:introduction}), therefore gamma-ray data compliant with those
can be directly read with the package. Version X.X of the \gadf specification
is supported in this release version. As both the \gadf data model and \gammapy
were initially conceived for \iact data analysis, input data are typically
grouped in \code{Observations}, corresponding to stable periods of data
aquisitions (typically $20 - 30\,{\rm min}$). A \code{DataStore} object,
gathering a collection of observations, can be created providing ancillary
files containing information about the telescope observation mode and the
content of the data unit of each file. The \code{DataStore} allows for
selecting a list of observations based on specific filters. As an illustrative
example, Fig.~\ref{fig*:minted:gp_data} shows how to create a
\code{DataStore} and how to obtain the observations corresponding to given
identification numbers (IDs).

The so-called DL3 files represented by the \code{Observation} class consist
of two types of elements: a list of gamma-ray events with relevant physical
quantities for the successive analysis (estimated energy, direction and arrival
times) that is handled by the \code{EventList} class and an instrument
response function (IRF), providing the response of the system, typically
factorised in independent components (see the description in
Sec.~\ref{ssec:gammapy-irf}). The separate handling of event lists and IRFs
additionally allows for data from other gamma-ray instruments to be read. For
example, to read \fermi data, the user can read separately their event list
(already compliant with the \gadf specifications) and then find the appropriate
IRF class representing the response functions provided by \fermi, see
Sec.~\ref{ssec:fermi}.