### 2022-07-08 src/text/2-package-subsections/maps.tex

\subsection{gammapy.maps}
\label{ssec:gammapy-maps}
\todo{Laura Olivera-Nieto}

\begin{figure}
	\import{code-examples/generated/}{gp_maps}

	\caption{Using \code{gammapy.maps} to create a WCS, a HEALPix and a region
		based data structures. The initialisation parameters include
        consistently the positions of the center of the map, the pixel
        size, the extend of the map as well as the energy axis definition.
        } \label{ig*:minted:gp_maps}
\end{figure}

The \code{gammapy.maps} sub-package provides classes that represent data
structures associated with a set of coordinates or a region on a sphere. In
addition it allows to handle an arbitrary number of non-spatial data
dimensions, such as time or energy. It is organized around three types of
structures: geometries, sky-maps and map axes, which inherit from the base
classes \code{Geom}, \code{Map} and \code{MapAxis} respectively.

The geometry object defines the pixelization scheme and map boundaries. It also
provides methods to transform between sky and pixel coordinates. Maps consist
of a geometry instance together with a Numpy data array containing the corresponding
map values. Map axes contain a sequence of ordered values which define bins on
a given dimension, spatial or not. Map axes can have physical units attached to
them, as well as define non-linear spaced bins.

The sub-package provides a uniform API for the FITS World Coordinate System
(WCS)~\citep{Calabretta2002}, the HEALPix pixelization~\citep{Gorski2005} and
region-based data structure (see Figure~\ref{ig*:minted:gp_maps}).

% itemize because it helps me write, could also just be paragraphs...
\paragraph{WCS:}
The FITS WCS pixelization supports a different
number of projections to represent celestial spherical coordinates in a regular
rectangular grid. Gammapy provides full support to data structures using this
pixelization scheme.

\paragraph{HEALPix:}
This pixelization provides a
subdivision of a sphere in which each pixel covers the same surface area as
every other pixel. As a consequence, however, pixel shapes are no longer
rectangular, or regular. \gammapy provides limited support to HEALPix-based
maps, relying in some cases on projections to a local WCS grid.

\paragraph{Region:} In this case, instead of a fine spatial grid
dividing a rectangular sky region, the spatial dimension is reduced to a single
bin with an arbitrary shape, describing a region in the sky with that same
shape. Typically they are is used together with a non-spatial dimension, for
example an energy axis, to represent how a quantity varies in that dimension
inside the corresponding region.

Additionally, the \code{MapAxis} class provides a uniform API for axes representing
bins on any physical quantity, such as energy or angular offset. The special
case of time is covered by the dedicated \code{TimeMapAxis}, which allows time bins to
be non-contiguous, as it is often the case with observation time-stamps. The
generic class \code{LabelMapAxis} allows the creation of axes for non-numeric
entries.