### 2022-07-25 src/text/3-applications-subsections/sed-modelling-analysis.tex
Update sed modeling section

\subsection{Broadband SED Modeling}
\label{ssec:broadband-sed-modeling}
Figure \ref{fig:multi_instrument_analysis}. Add explanation text.