### 2022-07-25 src/text/3-applications-subsections/temporal-analysis.tex
Update figures

\subsection{Temporal analysis}
\label{ssec:temporal-analysis}
A common use case in most astrophysical scenarios is to study the temporal
variability of the source. The most basic way to do this is to construct a
lightcurve, i.e., the flux of the source in each given time bin. In \gammapy, this
is done by using the \code{LightCurveEstimator} which fits the normalisation of the
source in each energy band per observation, keeping other parameters constant.
For custom time binning, an observation needs to be split into finer time bins using
the \code{Observation.select\_time} method. Figure~\ref{fig:hess_lightcurve_pks}
shows the lightcurve of the blazar PKS~2155-304 in different energy bands as
observed by the H.E.S.S. telescope during an exceptional flare on the night of
July 29 - 30, 2006~\cite{2009A&A...502..749A}. The data is available publicly
as a part of the HESS-DL3-DR1~\cite{HESS-DL3-DR1}, and shipped with
\verb"GAMMAPY_DATA". Each observation is first split into 10 min smaller
observations, and spectra extracted for each of these within a 0.11 deg radius
around the source. A \code{PowerLawSpectralModel} is fit to all the datasets, leading
to a reconstructed index of $3.54 \pm 0.02$. With this assumed spectral model
the \code{LightCurveEstimator} runs directly for two energy bands, 0.5 - 1.5
TeV, and 1.5 - 20 TeV, respectively.
%
\begin{figure*}[t]
    \sidecaption
	\includegraphics[width=0.6666\textwidth]{figures/hess_lightcurve_pks.pdf}
	\caption{
        Binned light-curves in two different energy bands differ for the source
        PKS~2155-304 in two energy bands, (500 GeV - 1.5 TeV, and 1.5 TeV to 20 TeV)
        as observed by the H.E.S.S. telescopes in 2006. The colored markers
        show the flux points in the different energy bands. The horizontal
        error illustrates the width of the time bin of 10~min. The vertical
        error bars show the associated assymetrical flux errors. The marker
        is set to the center of the time bin.
    }
    \label{fig:hess_lightcurve_pks}
\end{figure*}
%
The obtained flux points can be analytically modelled using the available, or
user-implemented temporal models. Alternatively, instead of  extracting a
lightcurve, it is also possible to directly fit temporal models to the reduced
datasets. By associating an appropriate \code{SkyModel}, consisting of both temporal
and spectral components, or using custom temporal models with spectroscopic
variability, to each dataset, a joint fitting across the datasets will directly
return the best fit temporal and spectral parameters.