### 2022-07-26 src/text/2-package-subsections/analysis.tex
Add instruments IRF plot

\subsection{gammapy.analysis}
\label{ssec:gammapy-analysis}
The \code{gammapy.analysis} sub-module provides a  high-level interface for the most
common use cases identified in \gammaray analyses. The classes and methods
included can be used in Python scripts, notebooks or as commands within IPython
sessions. The high-level user interface can also be used to automatise
workflows driven by parameters declared in a configuration file in YAML format.
This way a full analysis can be executed via a single command line taking the
configuration file as input.

The \code{Analysis} class has the responsibility of orchestrating of the workflow
defined in the configuration \code{AnalysisConfig} objects and triggering the execution of
the \code{AnalysisStep} classes that defines the common use cases identified. These
steps include the following: observations selection from the DataStore,  data
reduction, excess map computation, model fitting, flux-points estimation, and
light-curves production. The structure of the \code{Analysis} class allows to define
custom AnalysisStep, so ones could design and configure arbitrarily complex
wokflows. \todo{the latter is not true for v1.0..., include example YAML file?}