### 2022-07-28 src/text/2-package-subsections/visualisation.tex

\subsection{gammapy.visualization}
\label{ssec:gammapy-visualization}
The \code{gammapy.visualization} sub-package contains helper functions
for plotting and visualizing analysis results and \gammapy~data structures.
This includes for example the visualization of reflected background regions across
multiple observations or plotting large parameter correlation matrices of
\gammapy models. It also includes a helper class to split
wide field Galactic survey images across multiple panels to fit a standard
paper size.

The sub-package also provides matplotlib implementations of specific
colormaps for false color image representation. Those colormaps have
been used historically larger collaborations in the very high energy
domain (such as \milagro or \hess) as "trademark" colormaps.
While we explicitly discourage the use of those colormaps for publication
of new results, because they do not follow modern visualization
standards, such as linear brightness gradients and accessibility
for visually impaired people, we still consider the colormaps
useful for reproducibility of past results.